Geometrization of characters of tori over non-archimedean local fields

### 2013-08-15 shanghai.tex
Moving computers. Minor changes, mostly to Part 2.

%\input{introduction}
%In his celebrated thesis, John Tate writes 
%\begin{quotation}
%    {\it Concerning the characters of $k^*$ [the
%    multiplicative group of a $\mathfrak{p}$-adic field], the
%    situation is different from that of $k^+$.  First of all, we are
%    interested in continuous multiplicative maps $\alpha \mapsto
%    c(\alpha)$ of $k^*$ into the complex numbers, not only in the
%    bounded ones, and shall call such a map a quasi-character,
%    reserving the word ``character'' for the conventional character of
%    absolute value $1$. Secondly, we shall find no model for the group
%    of quasi-characters, or even for the group of characters, though
%    such a model would be of the utmost importance.}
%  \cite{Tate:thesis}*{\S 2.3}
%\end{quotation}
%\noindent
What is a quasicharacter sheaf?
Let $G$ be a smooth commutative group scheme locally of finite
type over a finite field $\Fq$; $G$ need not be reductive, connected, or of
finite type over $\Fq$. Fix a prime $\ell$ that does not divide
$q$, write $m : \GxG \to G$ for the multiplication morphism,
$\bG$ for the base change of $G$ to $\bFq$ and
$\Frob{G} : \bG \to \bG$ for the Frobenius automorphism.
A quasicharacter sheaf on $G$ is a triple $\cs{L}\ceq
(\gcs{L},\mu,\phi)$ where $\gcs{L}$ is a rank-$1$ $\ell$-adic local system on $\bG$ and $\mu: m^*
\gcs{L} \to \LxL$ and $\phi : \Frob{G}^*\gcs{L} \to \gcs{L}$ are isomorphisms;
$\mu$ and $\phi$ are required to satisfy natural conditions explained below.

@@

Throughout this paper, $G$ is a commutative group scheme,
smooth and locally of finite type over a finite field $\Fq$, and $m : \GxG\to G$ is its multiplication morphism.

We fix an algebraic closure $\bFq$ of $\Fq$ and write $\bG$ for the
commutative group scheme $G \times_{\Spec{\Fq}} \Spec{\bFq}$ over $\bFq$ obtained by base change from $k$. The multiplication morphism for $\bG$ will be denoted by $\bm$.

@@

\begin{definition}\label{def:QC}
A \emph{quasicharacter sheaf on $G$} is a triple
$\cs{L}\ceq (\gcs{L},\mu,\phi)$ where:
\begin{enumerate}
\labitem{(CS.0)}{CS.0} $\gcs{L}$ is a rank-$1$ $\ell$-adic local system on $\bG$;
\labitem{(CS.1)}{CS.1} $\mu: \bm^* \gcs{L} \to \LxL$
  is an isomorphism of sheaves on $\bGxG$ such that the following diagram commutes, 
  where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;
%
  \[
  \begin{tikzcd}[row sep=30]
  \bm_3^*\gcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gcs{L} \boxtimes \gcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gcs{L} \boxtimes \bm^* \gcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gcs{L} \tight{0}{\boxtimes}{0} \LxL\\
  \end{tikzcd}
  \]

%
\labitem{(CS.2)}{CS.2} $\phi : \Frob{G}^* \gcs{L} \to \gcs{L}$ is an
  isomorphism of constructible $\ell$-adic sheaves on $\bG$ compatible with $\mu$ in the sense that the following diagram commutes.
%
  \[
  \begin{tikzcd}[row sep=20]
  \Frob{\GxxG}^* \bm^* \gcs{L} \arrow{rr}{\Frob{\GxxG}^*\mu}
    && \Frob{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob{G}^* \gcs{L} \arrow[swap]{d}{\bm^* \phi}
    && \Frob{G}^*\gcs{L}\boxtimes \Frob{G}^*\gcs{L} \dar{F\boxtimes \phi} \arrow[equal]{u} \\
    \bm^*\gcs{L} \arrow{rr}{\mu}
    && \LxL\\
  \end{tikzcd}
  \]
\end{enumerate}
%In short, a quasicharacter sheaf on $G$  is a rank-$1$ Weil local system $(\gcs{L},\phi)$ on $G$ equipped with an isomorphism $m^* (\gcs{L},\phi) \to (\gcs{L},\phi)\boxtimes (\gcs{L},\phi)$. 
\end{definition}

@@

\begin{proof}
The only non-trivial part is essential surjectivity. Set $G \ceq G_1\times G_1$. Let $\cs{L}$ be a geochar on $G$. Define $f : G\to G\times G$ by $f(g_1,g_2) = (g_1,e_2,e_1,g_2)$. Observe that $m\circ f = \id_G$. 
Then the isomorphism $\mu : m^* \gcs{L} \to \gcs{L}\boxtimes \gcs{L}$ gives an isomorphism
\[
f^*\mu : f^* m^* \gcs{L} \to f^*(\gcs{L}\boxtimes \gcs{L}) = f^*p_1^*\gcs{L}\otimes f^* p_2^*\gcs{L}.
\]
Observe that $p_1\circ f = r_1 \circ q_1$ and $f_2 = r_2\circ f$ where $p_1 , p_2 : G\times G \to G$ are the usual things and $r_1 : G_1\times G_2 \to G_1$ and $r_2 : G_1\times G_2 \to G_2$ are the obvious things and $q_1 : G_1\to G_1\times G_2$ and  $q_2 : G_2\to G_1\times G_2$ are the usual things. Then $f^*p_1^*\gcs{L} = r_1^* q_1^* \gcs{L}$ and $f^*p_2^*\gcs{L} = r_2^* q_2^* \gcs{L}$. Now $\cs{L}_1 \ceq r_1^* \cs{L}$ is a geochar on $G_1$ and $\cs{L}_2 \ceq r_2^* \cs{L}$ is a geochar on $G_2$, by Proposition~\ref{prop:pullback}. Then
$ f^*p_1^*\gcs{L}\otimes f^* p_2^*\gcs{L} = r_1^*\gcs{L}_1\otimes r_2^* \gcs{L}_2 = \gcs{L}_1\boxtimes \gcs{L}_2$. In this way, $\mu$ determines an isomorphism
\[
f^*\mu : \gcs{L} \to  \gcs{L}_1\boxtimes \gcs{L}_2.
\]
So $\boxtimes$ is essentially surjective.
\end{proof}

@@

\begin{definition}
Let $\QCb(G)$ be the category of pairs $(\cs{L}_0,\mu_0)$ 
where $\cs{L}_0$ a rank-$1$ $\ell$-adic local system on $G$, 
equipped with an isomorphism $\mu_0 : m^* \cs{L}_0 \to \cs{L}_0 \boxtimes \cs{L}_0$ 
satisfying the analogue of Condition~\ref{CS.1} on $G$; 
morphisms in $\QCb(G)$ are defined as in (the second part of) Condition~\ref{CS.2}. 
This is the category of \emph{bounded quasicharacter sheaves} on $G$. 
\end{definition}

@@

Let us say that \cdef{global section}
of a quasicharacter sheaf $\cs{L}$ on an \'etale group scheme $G$ as a choice

@@

\begin{proof}
  Because $G$ is smooth over $\Fq$, the identity component, 
  $G^0$, is a connected algebraic group over $\Fq$, 
  clopen in $G$ \cite{vdG&M}*{3.17}. 
  By Proposition~\ref{prop:connected}, every
  quasicharacter sheaf on $G^0$ is isomorphic to a 
  finite quasicharacter sheaf on $G^0$, 
  so to prove the proposition it suffices to show that every 
  finite geochar on $G^0$ extends to a geochar on $G$.

 Let $(\pi,\psi)$ be a finite geochar on $G^0$.
 By Lemma~\ref{lemma:ext}, there is an extension of the 
 descrete isogeny $\pi : B \to G^0$ 
 (see Section~\ref{sec:finite} for the definition of a discrete isogeny) 
 to a discrete isogeny $f : H \to G^0$ 
 such that $\pi_0(f) : \pi_0(H)\to \pi_0(G)$ is an isomorphism.
 Then $(f,\psi)$ is a finite geochar on $G$.
 Then $(f,\psi)\vert_{G^0} \iso (\pi,\psi)$,
 completing the proof that $\QC(G)\to \QC(G^0)$ is essentially surjective.
\end{proof}

\begin{lemma}\label{lemma:ext}
Every discrete isogeny onto $G^0$ extends to a discrete isogeny onto $G$ such that the component group of the covering space is isomorphic to the component group of $G$.
\end{lemma}

@@

We saw, in Proposition~\ref{prop:functorialG}, that the trace of Frobenius  $\trFrob{G} : \QCiso{G} \to G(\Fq)^*$ is a functorial group homomorphism. Our main result regarding quasicharacter sheaves shows that trace of Frobenius is an isomorphism in the case that we care about.

%\input{snake}
\begin{theorem}\label{thm:snake}
  If $G$ is a smooth abelian group scheme, locally of finite type over
  $\Fq$, and the abelian group $\pi_0(G)(\bFq)$ (the geometric
  component group) is finitely generated, then the trace of Frobenius
  \[
  \trFrob{G} : \QCiso{G} \to G(\Fq)^*
  \]
  is an isomorphism of groups.
\end{theorem}

@@

With Theorem~\ref{thm:snake} on hand, 
we can see why neither bounded quasicharacter sheaves 
nor finite quasicharacter sheaves are up to the task of 
geometrizing characters of $G(\Fq)$ in the generality we need.  
%
Consider the case when $G$ is the discrete etale group scheme $\ZZ$. 
Let $\chi : G(\Fq) \to \EEx$ be the character determined by $\chi(1) = \ell$ 
and let $\cs{L}$ be a quasicharacter sheaf in the isomorphism class 
corresponding to the character $\chi : G(\Fq) \to \EEx$ under Theorem~\ref{thm:snake}. 
Then $\trFrob{\cs{L}}(1) = \ell$. Suppose $\cs{L}$ was a bounded quasicharacter sheaf. 
Then the restriction of $\cs{L}$ to the connected component $1\in G$ 
would be a rank-1 $\ell$-adic local system on $1$, 
thus an $\ell$-adic character $\rho_1 : \pi_1(1,{\bar 1}) \to \EEx$ with $\rho_1(\Frob{\Fq}) = \ell$. But $\pi_1(1,{\bar 1}) = \Gal(\bFq/\Fq) \iso {\hat \ZZ}$ is compact, 
while $\rho_1(\pi_1(1,{\bar 1}))$ is not compact in $\EEx$ (it contains $\ell^\ZZ$). 
So the Weil sheaf $\gcs{L}$ does not descend to a local system on $G$.
It follows that the quasicharacter sheaf $\cs{L}$ is not a bounded quasicharacter sheaf.
%
We can play the same game with the character $\chi : G(\Fq) \to \EEx$ 
defined by $\chi(1) = 1+\ell$. In this case, the corresponding quasicharacter sheaf $\cs{L}$ is bounded but not finite, since $1+\ell$ is not a root of unity. 
%
These simple examples illustrate why the categories $\QCb(G)$ and $\QCf(G)$ are inadequate to geometrize characters of $G(\Fq)$.

@@

Let $K$ be a non-archimedean local field with residue field $\Fq$.
We place no restriction on the characteristic of $K$. 
We write $\OK$ for the ring of integers of $K$ and $\pK$ for the maximal ideal in $\OK$. 
For any positive integer $d$, set $A_d\ceq \OK/\pK^d$.
Let $T$ be an algebraic torus over $K$.
We write $\Hom_{< d}(T(K),\EEx)$ for the group of characters of $T(K)$ depth less than $d$ 
and $\Hom_{<\infty}(T(K),\EEx)$ for the group of characters of $T(K)$ of arbitrary finite depth.

@@

\section{The geometrization of quasicharacters} \label{sec:quasichar}

@@

\begin{corollary}\label{cor:application}
\begin{enumerate}
\item 
For every algebraic torus $T$ over $K$, the trace of Frobenius 
  \[
\trFrob{T}:  \QCiso{\GN{T}} \to \Hom_{\text{qc}}(T(K),\EEx)
  \]
is an isomomorphism. 
\item The isomorphism $\trFrob{T}$ is functorial in the following sense: $T \mapsto \trFrob{T}$ is a natural transformation
from the additive functor
\[T \mapsto \QCiso{\GN{T}}\]
to the additive functor
\[
T \mapsto \Hom_{\text{qc}}(T(K),\EEx).
\]
\end{enumerate}
\end{corollary}

@@

We recall the notion of $N$-congruent tori from \cite{CY}*{\S 2}.  
Suppose $T$ and $T'$ are tori over non-archimedian local fields $K$ and $K'$, 
splitting over $L$ and $L'$ respectively.
Then $T$ and $T'$ are said to be \emph{$N$-congruent} if there are isomorphisms
 \begin{align*}
  \alpha : \OL/\pi_K^N \OL &\to \OO{L'}/\pi_{K'}^N \OO{L'} \\
  \beta : \Gal(L/K) &\to \Gal(L'/K') \\
  \phi : X^*(T) &\to X^*(T')
 \end{align*}
 satisfying the following conditions:
 \begin{enumerate}
  \item $\alpha$ induces an isomorphism $\OK/\pi_K^N \OK \to \OO{K'}/\pi_{K'}^N \OO{K'}$,
  \item $\alpha$ is $\Gal(L/K)$-equivariant relative to $\beta$: $\alpha(\gamma x) = \beta(\gamma) \alpha(x)$
  for $\gamma \in \Gal(L/K)$ and $x \in \OL/\pi_K^N \OL$,
  \item $\phi$ is $\Gal(L/K)$-equivariant relative to $\beta$.
 \end{enumerate}
It follows from this definition that if $T$ and $T'$ are $N$-congruent then $\alpha$, $\beta$ and $\phi$ determine an isomorphism
\begin{equation}\label{transfer}
  \Hom_{<N}(T(K), \EEx) \iso \Hom_{<N}(T'(K'),\EEx).
\end{equation}
Note that if $T$ and $T'$ are $N$-congruent, then they are $N'$ congruent for every $N' \leq N$.
In this section we ask if the isomorphism \eqref{transfer} comes from an equivalence of categories of quasicharacter sheaves on $\GN{T}_d$ and those on $\GN{T'}_d$.

One of the main results of \cite{CY} gives an isomorphism of group schemes between
$\TT_d$ and $\TT'_d$ assuming that $T$ and $T'$ are sufficiently congruent.
They define a quantity $h$ as the smallest integer so that $\pi^h$ lies in the
Jacobian ideal associated to a natural embedding of $T$ into an induced torus \cite{CY}*{\S 8.1}; let $h'$ be this quantity for $T'$.
Then it follows from \cite{CY}*{Thm. 8.5} that if $T$ and $T'$ are $N$-congruent and $h=h'$ and $N > 3h$, then there is a canonical isomorphism
 \[
  \TT_{N-3h} \to \TT'_{N-3h}
 \]
 determined by $\alpha, \beta$ and $\phi$.
As an immediate consequence of this theorem and Theorem \ref{thm:application} we have the following

\begin{proposition}
 With notation above, suppose that $T$ and $T'$ are $N$-congruent and $h=h'$ and $N > 3h$.  Set $d = N-3h$.
 Then there is a canonical equivalence of categories
 \[
  \QC(\GN{T}'_{d}) \to \QC(\GN{T}_{d})
 \]
 determined by $\alpha, \beta$ and $\phi$ inducing the isomorphism
 \[
  \Hom_{<d}(T'(K'), \EEx) \to \Hom_{<d}(T(K), \EEx).
 \]
\end{proposition}

@@

As indicated in the Introduction, we are tantalized 
by the idea that quasicharacter sheaves for $T$ 
may play a key role in a geometric, categorial version 
of the local Langlands correspondence for algebraic tori 
over non-archimedean local fields. 
As with the local Langlands correspondence for algebraic tori, 
it will be possible to reduce the essential points of the argument to the case $T= \Gm{K}$, provided that the category of quasicharacter sheaves behaves properly with regard to Weil restriction; see \cite{Yu:tori}*{Thm 7.5 (2)}. 
The next proposition shows that the category of quasicharacter sheaves for 
algebraic tori does indeed behave properly with regard to Weil restriction.

@@

\bib{Tate:thesis}{article}{
   author={Tate, J. T.},
   title={Fourier analysis in number fields, and Hecke's zeta-functions},
   conference={
      title={Algebraic Number Theory (Proc. Instructional Conf., Brighton,
      1965)},
   },
   book={
      publisher={Thompson, Washington, D.C.},
   },
   date={1967},
   pages={305--347},
   review={\MR{0217026 (36 \#121)}},
}

>>>>

%\input{introduction}
%In his celebrated thesis, John Tate writes 
%\begin{quotation}
%    {\it Concerning the characters of $k^*$ [the
%    multiplicative group of a $\mathfrak{p}$-adic field], the
%    situation is different from that of $k^+$.  First of all, we are
%    interested in continuous multiplicative maps $\alpha \mapsto
%    c(\alpha)$ of $k^*$ into the complex numbers, not only in the
%    bounded ones, and shall call such a map a quasi-character,
%    reserving the word ``character'' for the conventional character of
%    absolute value $1$. Secondly, we shall find no model for the group
%    of quasi-characters, or even for the group of characters, though
%    such a model would be of the utmost importance.}
%  \cite{Tate:thesis}*{\S 2.3}
%\end{quotation}
%\noindent
What is a quasicharacter sheaf?
Let $G$ be a smooth commutative group scheme locally of finite
type over a finite field $\Fq$; $G$ need not be reductive, connected, or of
finite type over $\Fq$. Fix a prime $\ell$ invertible in $\Fq$
and write $m : \GxG \to G$ for the multiplication morphism,
$\bG$ for the base change of $G$ to $\bFq$ and
$\Frob{G} : \bG \to \bG$ for the Frobenius automorphism.
A quasicharacter sheaf on $G$ is a 
\Clifton{ [Use this?:] . . . is an $\ell$-adic Weil local system of rank-$1$ on $G$, equipped with an isomorphism ... }
triple $\cs{L}\ceq
(\gcs{L},\mu,\phi)$ where $\gcs{L}$ is an $\ell$-adic local system on $\bG$ of rank-1 and $\mu: m^*
\gcs{L} \to \LxL$ and $\phi : \Frob{G}^*\gcs{L} \to \gcs{L}$ are isomorphisms;
$\mu$ and $\phi$ are required to satisfy natural conditions explained below.

@@

Throughout this paper, $G$ is a smooth commutative group scheme locally of finite type over a finite field $\Fq$, and $m : \GxG\to G$ is its multiplication morphism.

We fix an algebraic closure $\bFq$ of $\Fq$ and write $\bG$ for the
smooth commutative group scheme $G \times_{\Spec{\Fq}} \Spec{\bFq}$ over $\bFq$ obtained by base change from $k$. The multiplication morphism for $\bG$ will be denoted by $\bm$.

@@

\begin{definition}\label{def:QC}
A \emph{quasicharacter sheaf on $G$} is a triple
$\cs{L}\ceq (\gcs{L},\mu,\phi)$ where:
\begin{enumerate}
\labitem{(CS.0)}{CS.0} $\gcs{L}$ is an $\ell$-adic local system on $\bG$ (a constructible $\ell$-adic sheaf on $\bG$, locally constant on each connected component) of rank-1; 
\labitem{(CS.1)}{CS.1} $\mu: \bm^* \gcs{L} \to \LxL$
  is an isomorphism of sheaves on $\bGxG$ such that the following diagram commutes, 
  where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;
%
  \[
  \begin{tikzcd}[row sep=30]
  \bm_3^*\gcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gcs{L} \boxtimes \gcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gcs{L} \boxtimes \bm^* \gcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gcs{L} \tight{0}{\boxtimes}{0} \LxL\\
  \end{tikzcd}
  \]

%
\labitem{(CS.2)}{CS.2} $\phi : \Frob{G}^* \gcs{L} \to \gcs{L}$ is an
  isomorphism of constructible $\ell$-adic sheaves on $\bG$ compatible with $\mu$ in the sense that the following diagram commutes.
%
  \[
  \begin{tikzcd}[row sep=20]
  \Frob{\GxxG}^* \bm^* \gcs{L} \arrow{rr}{\Frob{\GxxG}^*\mu}
    && \Frob{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob{G}^* \gcs{L} \arrow[swap]{d}{\bm^* \phi}
    && \Frob{G}^*\gcs{L}\boxtimes \Frob{G}^*\gcs{L} \dar{F\boxtimes \phi} \arrow[equal]{u} \\
    \bm^*\gcs{L} \arrow{rr}{\mu}
    && \LxL\\
  \end{tikzcd}
  \]
\end{enumerate}
%In short, a quasicharacter sheaf on $G$  is a rank-1 Weil local system $(\gcs{L},\phi)$ on $G$ equipped with an isomorphism $m^* (\gcs{L},\phi) \to (\gcs{L},\phi)\boxtimes (\gcs{L},\phi)$. 
\end{definition}

@@

\begin{proof}
The only non-trivial part is essential surjectivity. Set $G \ceq G_1\times G_1$. Let $\cs{L}$ be a quasicharacter sheaf on $G$. Define $f : G\to G\times G$ by $f(g_1,g_2) = (g_1,e_2,e_1,g_2)$. Observe that $m\circ f = \id_G$. 
Then the isomorphism $\mu : m^* \gcs{L} \to \gcs{L}\boxtimes \gcs{L}$ gives an isomorphism
\[
f^*\mu : f^* m^* \gcs{L} \to f^*(\gcs{L}\boxtimes \gcs{L}) = f^*p_1^*\gcs{L}\otimes f^* p_2^*\gcs{L}.
\]
Observe that $p_1\circ f = r_1 \circ q_1$ and $f_2 = r_2\circ f$ where $p_1 , p_2 : G\times G \to G$ are the usual things and $r_1 : G_1\times G_2 \to G_1$ and $r_2 : G_1\times G_2 \to G_2$ are the obvious things and $q_1 : G_1\to G_1\times G_2$ and  $q_2 : G_2\to G_1\times G_2$ are the usual things. Then $f^*p_1^*\gcs{L} = r_1^* q_1^* \gcs{L}$ and $f^*p_2^*\gcs{L} = r_2^* q_2^* \gcs{L}$. Now $\cs{L}_1 \ceq r_1^* \cs{L}$ is a quasicharacter sheaf on $G_1$ and $\cs{L}_2 \ceq r_2^* \cs{L}$ is a quasicharacter sheaf on $G_2$, by Proposition~\ref{prop:pullback}. Then
$ f^*p_1^*\gcs{L}\otimes f^* p_2^*\gcs{L} = r_1^*\gcs{L}_1\otimes r_2^* \gcs{L}_2 = \gcs{L}_1\boxtimes \gcs{L}_2$. In this way, $\mu$ determines an isomorphism
\[
f^*\mu : \gcs{L} \to  \gcs{L}_1\boxtimes \gcs{L}_2.
\]
So $\boxtimes$ is essentially surjective.
\end{proof}

@@

\begin{definition}
Let $\QCb(G)$ be the category of pairs $(\cs{L}_0,\mu_0)$ 
where $\cs{L}_0$ an $\ell$-adic local system on $G$ of rank-1, 
equipped with an isomorphism $\mu_0 : m^* \cs{L}_0 \to \cs{L}_0 \boxtimes \cs{L}_0$ 
satisfying the analogue of Condition~\ref{CS.1} on $G$; 
morphisms in $\QCb(G)$ are defined as in (the second part of) Condition~\ref{CS.2}. 
This is the category of \emph{bounded quasicharacter sheaves} on $G$. 
\end{definition}

@@

@@

\begin{proof}
  Because $G$ is smooth over $\Fq$, the identity component, 
  $G^0$, is a connected algebraic group over $\Fq$, 
  clopen in $G$ \cite{vdG&M}*{3.17}. 
  By Proposition~\ref{prop:connected}, every
  quasicharacter sheaf on $G^0$ is isomorphic to a 
  finite quasicharacter sheaf on $G^0$, 
  so to prove the proposition it suffices to show that every 
  finite quasicharacter sheaf on $G^0$ extends to a quasicharacter sheaf on $G$.

 Let $(\pi,\psi)$ be a finite quasicharacter sheaf on $G^0$.
 By Lemma~\ref{lemma:ext}, there is an extension of the 
 discrete isogeny $\pi : B \to G^0$ 
 (see Section~\ref{sec:finite} for the definition of a discrete isogeny) 
 to a discrete isogeny $f : H \to G^0$ 
 such that $\pi_0(f) : \pi_0(H)\to \pi_0(G)$ is an isomorphism.
 Then $(f,\psi)$ is a finite quasicharacter sheaf on $G$.
 Then $(f,\psi)\vert_{G^0} \iso (\pi,\psi)$,
 completing the proof that $\QC(G)\to \QC(G^0)$ is essentially surjective.
\end{proof}

\begin{lemma}\label{lemma:ext}
Every discrete isogeny (see Section~\ref{sec:finite}) to $G^0$ extends to a discrete isogeny to $G$ such that the component group of the covering space is isomorphic to the component group of $G$.
\end{lemma}

@@

We saw, in Proposition~\ref{prop:functorialG}, that the trace of Frobenius  $\trFrob{G} : \QCiso{G} \to G(\Fq)^*$ is a functorial group homomorphism. Our main result regarding quasicharacter sheaves, below, shows that trace of Frobenius is an isomorphism in the case that we care about.

%\input{snake}
\begin{theorem}\label{thm:snake}
  If $G$ is a smooth abelian group scheme, locally of finite type over
  $\Fq$, and if the abelian group $\pi_0(G)(\bFq)$ (the geometric
  component group) is finitely generated, then the trace of Frobenius
  \[
  \trFrob{G} : \QCiso{G} \to G(\Fq)^*
  \]
  is an isomorphism of groups.
\end{theorem}

@@

With Theorem~\ref{thm:snake} on hand, 
we can see why neither bounded quasicharacter sheaves 
nor finite quasicharacter sheaves are up to the task of 
geometrizing characters of $G(\Fq)$ in the generality we need.  
%
Consider the case when $G$ is the discrete etale group scheme $\ZZ$. 
Let $\chi : G(\Fq) \to \EEx$ be the character determined by $\chi(1) = \ell$ 
and let $\cs{L}$ be a quasicharacter sheaf in the isomorphism class 
corresponding to the character $\chi : G(\Fq) \to \EEx$ under Theorem~\ref{thm:snake}. 
Then $\trFrob{\cs{L}}(1) = \ell$. Suppose $\cs{L}$ was a bounded quasicharacter sheaf. 
Then the restriction of $\cs{L}$ to the connected component $1\in G$ 
would be a $\ell$-adic local system on $1$ of rank-1, 
thus an $\ell$-adic character $\rho_1 : \pi_1(1,{\bar 1}) \to \EEx$ with $\rho_1(\Frob{\Fq}) = \ell$. But $\pi_1(1,{\bar 1}) = \Gal(\bFq/\Fq) \iso {\hat \ZZ}$ is compact, 
while $\rho_1(\pi_1(1,{\bar 1}))$ is not compact in $\EEx$ (it contains $\ell^\ZZ$). 
So the Weil sheaf $\gcs{L}$ does not descend to a local system on $G$.
It follows that the quasicharacter sheaf $\cs{L}$ is not a bounded quasicharacter sheaf.
%
We can play the same game with the character $\chi : G(\Fq) \to \EEx$ 
defined by $\chi(1) = 1+\ell$. In this case, the corresponding quasicharacter sheaf $\cs{L}$ is bounded, defined at $1$ by a pro-etale cover of $1$, but not finite, since $1+\ell$ is not a root of unity. 
%
These simple examples illustrate why the categories $\QCb(G)$ and $\QCf(G)$ are inadequate to geometrize characters of $G(\Fq)$.

@@

Let $K$ be a non-archimedean local field with residue field $\Fq$.
We place no restriction on the characteristic of $K$. 
We write $\OK$ for the ring of integers of $K$ and $\pK$ for the maximal ideal in $\OK$. 
For any positive integer $d$, set $A_d\ceq \OK/\pK^d$.
Let $T$ be an algebraic torus over $K$.
We write $\Hom_{< d}(T(K),\EEx)$ for the group of characters of $T(K)$ depth less than $d$ 
and $\Hom_{<\infty}(T(K),\EEx)$ for the group of characters of $T(K)$ of arbitrary finite depth; one might call these quasicharacters of $T(K)$, expanding on \cite{Tate:thesis}*{2.3}.

@@

\section{The geometrization of characters of tori over local fields} \label{sec:quasichar}

@@

\begin{corollary}\label{cor:application}
\begin{enumerate}
\item 
For every algebraic torus $T$ over $K$, the trace of Frobenius 
  \[
\trFrob{T}:  \QCiso{\GN{T}} \to \Hom_{<\infty}(T(K),\EEx)
  \]
is an isomomorphism. 
\item The isomorphism $\trFrob{T}$ is functorial in the following sense: $T \mapsto \trFrob{T}$ is a natural transformation
from the additive functor
\[T \mapsto \QCiso{\GN{T}}\]
to the additive functor
\[
T \mapsto \Hom_{<\infty}(T(K),\EEx).
\]
\end{enumerate}
\end{corollary}

\begin{proof}
\Clifton{I'll say a few words here about the ind-category $\QC(\GN{T})$.}

\end{proof}

\Clifton{Make comments here about the depth-preserving isomorphism  . .  .}

@@

We recall the notion of $N$-congruent tori from \cite{CY}*{\S 2}.  
Suppose $T$ and $T'$ are tori over non-archimedian local fields $K$ and $K'$, 
splitting over $L$ and $L'$ respectively.
Then $T$ and $T'$ are said to be \emph{$N$-congruent} if there are isomorphisms
 \begin{align*}
  \alpha : \OL/\pi_K^N \OL &\to \OO{L'}/\pi_{K'}^N \OO{L'} \\
  \beta : \Gal(L/K) &\to \Gal(L'/K') \\
  \phi : X^*(T) &\to X^*(T')
 \end{align*}
 satisfying the conditions
 \begin{enumerate}
  \item $\alpha$ induces an isomorphism $\OK/\pi_K^N \OK \to \OO{K'}/\pi_{K'}^N \OO{K'}$,
  \item $\alpha$ is $\Gal(L/K)$-equivariant relative to $\beta$, and
  %: $\alpha(\gamma x) = \beta(\gamma) \alpha(x)$ for $\gamma \in \Gal(L/K)$ and $x \in \OL/\pi_K^N \OL$,
  \item $\phi$ is $\Gal(L/K)$-equivariant relative to $\beta$.
 \end{enumerate}
If $T$ and $T'$ are $N$-congruent then $\alpha$, $\beta$ and $\phi$ determine an isomorphism
\begin{equation}\label{transfer}
  \Hom_{<N}(T(K), \EEx) \iso \Hom_{<N}(T'(K'),\EEx).
\end{equation}
Note that if $T$ and $T'$ are $N$-congruent, then they are $N'$ congruent for every $N' \leq N$, the isomorphism \eqref{transfer} is depth-preserving at integers less than $N$.
In this section we ask if the isomorphism \eqref{transfer} comes from an equivalence of categories of quasicharacter sheaves on $\GN{T}_d$ and those on $\GN{T}'_d$.

One of the main results of \cite{CY} gives an isomorphism of group schemes between
$\TT_d$ and $\TT'_d$ assuming that $T$ and $T'$ are sufficiently congruent.
They define a quantity $h$ as the smallest integer so that $\pi^h$ lies in the
Jacobian ideal associated to a natural embedding of $T$ into an induced torus \cite{CY}*{\S 8.1}; let $h'$ be this quantity for $T'$.
Then it follows from \cite{CY}*{Thm. 8.5} that if $T$ and $T'$ are $N$-congruent and $h=h'$ and $N > 3h$, 
\Clifton{Is this correct?}
then there is a canonical isomorphism of smooth group schemes
 \[
  \TT_{N-3h} \to \TT'_{N-3h}
 \]
 determined by $\alpha, \beta$ and $\phi$.
Combining this with Theorem \ref{thm:application} gives the geometrization and categorification of the transfer of characters.

\begin{corollary}\label{cor:transfer}
 With notation above, suppose that $T$ and $T'$ are $N$-congruent and $h=h'$ and $N > 3h$.  Set $d = N-3h$.
 Then there is a canonical equivalence of categories
 \[
  \QC(\GN{T}'_{d}) \to \QC(\GN{T}_{d})
 \]
 determined by $\alpha, \beta$ and $\phi$ inducing the isomorphism
 \[
  \Hom_{<d}(T'(K'), \EEx) \to \Hom_{<d}(T(K), \EEx)
 \]
 through the trace of Frobenius.
\end{corollary}

@@

As indicated in the Introduction, we are tantalized 
by the idea that quasicharacter sheaves for $T$ 
may play a key role in a geometric, categorial version 
of the local Langlands correspondence for algebraic tori 
over non-archimedean local fields. 
As with the local Langlands correspondence for algebraic tori over local fields, 
it will be possible to reduce the essential points of the argument to the case $T= \Gm{K}$, provided that the category of quasicharacter sheaves behaves properly with regard to Weil restriction; see \cite{Yu:tori}*{Thm 7.5 (2)}. 
The next proposition shows that the category of quasicharacter sheaves for 
algebraic tori does indeed behave properly with regard to Weil restriction.

@@

\bib{Tate:thesis}{article}{
   author={Tate, J. T.},
   title={Fourier analysis in number fields, and Hecke's zeta-functions},
   conference={
      title={Algebraic Number Theory (Proc. Instructional Conf., Brighton,
      1965)},
   },
   book={
      publisher={Thompson, Washington, D.C.},
   },
   date={1967},
   pages={305--347},
%   review={\MR{0217026 (36 \#121)}},
}
